---
title: "Atmosphärische Strahlung - Excercise 9"
author: "Johannes Röttenbacher"
date: "22.01.2019"
output:
  html_document:
    df_print: paged
  html_notebook: 
    fig_caption: yes
  pdf_document: 
    df_print: kable
    fig_caption: yes
    fig_height: 5.5
    fig_width: 8
lang: German
toc: null
---

# 9.1 Monte Carlo Model

### Python Code to run the Monte Carlo Model

The idea is to have a function which runs for each layer the photon needs to pass through. Then the cloud is an additional function. All photons are counted and split into groups depending on where and how they leave the system. Photons can be either reflected $(F^\uparrow_{TOA})$, transmitted to the ground $(F^\downarrow_{BOA})$ or absorbed. The photons which reach the ground and are not reflected by the surface are divided into diffuse $(F^\downarrow_{BOA, diff})$and direct $(F^\downarrow_{BOA, dir})$ irradiance by two methods. The first divides them depending on whether they have been scattered and the second on their deviation from the solar zenith angle.

```{python eval=FALSE, include=TRUE}
# -*- coding: utf-8 -*-
"""
Created on Mon Dec  25 10:46:25 2018

@author: Johannes Roettenbacher

Script for a Monte-Carlo simulation for calculating radiative transfer through a single cloud with variable number
of layers.
Goal in Task A: Calculate and plot F upward TOA, F downward BOA in dependence of optical thickness tau (C_ext),
single scattering albedo omega and scattering phase function represented by the asymmetry parameter g.
Goal in Task B: Differentiate FBOA into direct transmitted and scattered diffuse radiation
"""
# %%
import numpy as np
import pandas


# %% layer function


def layer(tau_layer, angle, fun_layer, fun_omega, fun_g):
    """
    Computes whether a photon is extinct or not and whether it is scattered or absorbed.
    If it is scattered the scatter direction is computed.

    Parameters
    ----
    tau_layer: float
        thickness of layer
    angle: float
        incoming angle of photon in radiant
    fun_layer: float
        layer photon is currently in. 0 = ground
    fun_omega: float
        defines single scattering albedo
    fun_g: float
        defines asymmetry parameter g
    Return
    ----
    angle: float
           the angle at which the photon continues its path through the cloud
    fun_layer: integer
               the layer the photon is in after the processes
    scattered: bool
               variable which tells whether photon has been scattered or not
    """
    tau1 = abs(np.cos(angle) * np.log(np.random.random()))
    if tau1 < tau_layer:
        scattered = True
        if np.random.random() < fun_omega:
            my_scatter = (0.5 / fun_g) * ((1 + fun_g ** 2) - ((1 - fun_g ** 2) / (1 + fun_g - 2 * fun_g * np.random.random())) ** 2)
            angle = angle + np.arccos(my_scatter)
            if angle > (2 * np.pi):
                angle = angle - 2 * np.pi
        else:
            fun_layer = None
    else:
        scattered = False
    if angle < (np.pi / 2) or angle > (3 * np.pi / 2):
        direction = -1
    else:
        direction = 1
    if fun_layer is not None:
        fun_layer = fun_layer + direction
    return angle, fun_layer, scattered


# %% test what layer thickness is appropriate

surface_albedo = 0.05
omega = 1  # single scattering albedo
g = 0.9  # asymmetry parameter
tau = 10  # cloud optical thickness
n_layer = np.array(range(10, 101))
n_photon = 5000
delta_tau = (tau / n_layer)  # optical thickness of each layer
output = np.empty((len(n_layer), 3))
for i in range(len(n_layer)):
    print(i / len(n_layer))
    ground_count = 0  # counter for photons reaching the ground
    toa_up_count = 0  # counter for photons reflected into space
    in_cloud_count = 0  # counter for photons absorbed in the cloud
    for n_p in range(1, n_photon + 1):
        theta = 20 / 180 * np.pi
        photon_in_cloud = True
        in_layer = n_layer[i]
        while photon_in_cloud is True:
            if in_layer == 0:
                if np.random.random() > surface_albedo:
                    photon_in_cloud = False
                    ground_count += 1
                else:
                    in_layer += 1
            if in_layer is None:
                photon_in_cloud = False
                in_cloud_count += 1
            theta, in_layer, scattered = layer(tau_layer=delta_tau[i], angle=theta, fun_layer=in_layer,
                                               fun_omega=1, fun_g=0.9)
            if in_layer == n_layer[i] + 1:
                photon_in_cloud = False
                toa_up_count += 1
    output[i] = (n_layer[i], toa_up_count / n_photon, ground_count / n_photon)

df_output = pandas.DataFrame({"n_layer": output[:, 0], "FTOA": output[:, 1], "FBOA": output[:, 2]})
df_output.to_csv(
    "C:/Users/Johannes/Nextcloud/Studium/Atmosphaerische_Strahlung/Atmo_Strahlung/delta_tau_layer_depend.csv",
    index=False)


# %% monte_carlo method


def monte_carlo(tau, omega, g, n_photon=5000, theta_in=20, delta_tau=0.2):
    """
    Runs a MonteCarlo simulation with the given parameters.
    Parameters
    ----
    tau: integer
        optical thickness of cloud
    omega: float
        single scattering albedo
    g: float
        asymmetry parameter of the phase function
    n_photon: integer
        number of photons with which the model should run
    theta_in: float
        Incoming solar zenith angle for photon hitting cloud top in degree
    delta_tau: float
        thickness of each individual cloud layer, default = 0.2
    Returns
    ----
    Input parameters tau, omega, g and theta_in. Additionally the fraction of FTOA, FBOA,
    FBOA_diffuse1/2, FBOA_direct1/2 and in cloud absorbed photons is returned.
    """
    ground_count = 0  # counter for photons reaching the ground
    toa_up_count = 0  # counter for photons reflected into space
    in_cloud_count = 0  # counter for photons absorbed in the cloud
    boa_direct1 = 0  # counter for photons directly transmitted to surface, method 1
    boa_diffuse1 = 0  # counter for photons diffusely transmitted to surface, method 1
    boa_direct2 = 0  # counter for photons directly transmitted to surface, method 2
    boa_diffuse2 = 0  # counter for photons diffusely transmitted to surface, method 2
    n_layer = tau / delta_tau  # number of layers in cloud
    surface_albedo = 0.05
    for n_p in range(1, n_photon + 1):
        theta = theta_in / 180 * np.pi
        photon_in_cloud = True
        in_layer = n_layer
        diffuse = False
        while photon_in_cloud is True:
            if in_layer == 0:
                if np.random.random() > surface_albedo:
                    photon_in_cloud = False
                    ground_count += 1
                    if diffuse is True:
                        boa_diffuse1 += 1
                    else:
                        boa_direct1 += 1
                    if (theta_in - (1 / 180 * np.pi)) < theta < (theta_in + (1 / 180 * np.pi)):
                        boa_direct2 += 1
                    else:
                        boa_diffuse2 += 1
                else:
                    theta = theta + np.pi
                    if theta > (2 * np.pi):
                        theta = theta - (2 * np.pi)
                    in_layer += 1
            if in_layer is None:
                photon_in_cloud = False
                in_cloud_count += 1
            theta, in_layer, scattered = layer(tau_layer=delta_tau, angle=theta, fun_layer=in_layer, fun_omega=omega,
                                               fun_g=g)
            if scattered is True:
                diffuse = True
            if in_layer == n_layer + 1:
                photon_in_cloud = False
                toa_up_count += 1
    return tau, omega, g, toa_up_count / n_photon, ground_count / n_photon, in_cloud_count / n_photon, \
           boa_diffuse1 / (ground_count + 1), boa_direct1 / (ground_count + 1), \
           boa_diffuse2 / (ground_count + 1), boa_direct2 / (ground_count + 1), theta_in


# %% vary number of photons
n_photon = np.arange(1, 20002, 5000)
plot_data1 = np.empty((len(n_photon), 11))
h = 0
for x in range(len(n_photon)):
    plot_data1[h] = monte_carlo(tau=10, omega=0.999, g=0.9, n_photon=n_photon[x])
    h += 1
    print(h / (len(n_photon)))
df = pandas.DataFrame({"tau": plot_data1[:, 0], "omega": plot_data1[:, 1], "g": plot_data1[:, 2],
                       "FTOA": plot_data1[:, 3], "FBOA": plot_data1[:, 4], "Absorbed": plot_data1[:, 5]})
df.to_csv("C:/Users/Johannes/Nextcloud/Studium/Atmosphaerische_Strahlung/Atmo_Strahlung/ex09_plot_data1.csv",
          index=False)

# %% vary cloud optical thickness, single scattering albedo and asymmetry parameter
tau = np.array(range(1, 11))  # cloud optical thickness
omega = np.array([0.9, 0.99, 1])  # single scattering albedo
g = np.array([-0.9, -0.5, -0.01, 0.5, 0.9])  # asymmetry parameter
plot_data2 = np.empty((len(tau) * len(omega) * len(g), 11))
h = 0
for i in range(len(tau)):
    for j in range(len(omega)):
        for k in range(len(g)):
            plot_data2[h] = monte_carlo(tau=tau[i], omega=omega[j], g=g[k], n_photon=5000)
            h += 1
            print(h/(len(tau) * len(omega) * len(g)))

df1 = pandas.DataFrame({"tau": plot_data2[:, 0], "omega": plot_data2[:, 1], "g": plot_data2[:, 2],
                        "FTOA": plot_data2[:, 3], "FBOA": plot_data2[:, 4], "Absorbed": plot_data2[:, 5],
                        "BOA_diffuse1": plot_data2[:, 6], "BOA_direct1": plot_data2[:, 7],
                        "BOA_diffuse2": plot_data2[:, 8], "BOA_direct2": plot_data2[:, 9]})
df1.to_csv(
    "C:/Users/Johannes/Nextcloud/Studium/Atmosphaerische_Strahlung/Atmo_Strahlung/ex09_plot_data2_B.csv",
    index=False)

# %% vary cloud optical thickness, solar zenith angle and asymmetry parameter
tau = np.array(range(1, 11))  # cloud optical thickness
g = np.array([-0.9, -0.5, -0.01, 0.5, 0.9])  # asymmetry parameter
theta = np.arange(1, 82, 20)
plot_data3 = np.empty((len(tau) * len(theta) * len(g), 11))
h = 0
for i in range(len(tau)):
    for j in range(len(g)):
        for k in range(len(theta)):
            plot_data3[h] = monte_carlo(tau=tau[i], omega=1, g=g[j], n_photon=5000,
                                        theta_in=theta[k])
            h += 1
            print(h/(len(tau) * len(theta) * len(g)))

df2 = pandas.DataFrame({"tau": plot_data3[:, 0], "omega": plot_data3[:, 1], "g": plot_data3[:, 2],
                        "FTOA": plot_data3[:, 3], "FBOA": plot_data3[:, 4], "Absorbed": plot_data3[:, 5],
                        "BOA_diffuse1": plot_data3[:, 6], "BOA_direct1": plot_data3[:, 7],
                        "BOA_diffuse2": plot_data3[:, 8], "BOA_direct2": plot_data3[:, 9],
                        "theta": plot_data3[:, 10]})
df2.to_csv(
    "C:/Users/Johannes/Nextcloud/Studium/Atmosphaerische_Strahlung/Atmo_Strahlung/ex09_plot_data3_B.csv",
    index=False)


```


## Plotting the results

```{r Input from python, message=TRUE, warning=TRUE, include=FALSE}
library("tidyr")
delta_tau_layer_depend <- read.csv("./delta_tau_layer_depend.csv")
delta_tau_layer_depend_g <- gather(delta_tau_layer_depend, key = Class, value = Irradiance, FBOA, FTOA)

plot_data1 <- read.csv("./ex09_plot_data1.csv")
plot_data1$n_photon <- seq(from=1, to=20001, by=5000)
plot_data1 <- plot_data1[-1,]
plot_data1_g <- gather(plot_data1, key = Class, value = Irradiance, FTOA,
                       FBOA, Absorbed)
plot_data1_g$tau <- as.factor(plot_data1_g$tau)
plot_data1_g$omega <- as.factor(plot_data1_g$omega)
plot_data1_g$g <- as.factor(plot_data1_g$g)
plot_data1_g$n_photon <- as.factor(plot_data1_g$n_photon)
plot_data1_g$Class <- as.factor(plot_data1_g$Class)

plot_data2_B <- read.csv("./ex09_plot_data2_B.csv")
plot_data2_B_g <- gather(plot_data2_B, key = Class, value = Irradiance, FTOA, FBOA, Absorbed)
plot_data2_B_g$omega <- as.factor(plot_data2_B_g$omega)
plot_data2_B_g$g <- as.factor(plot_data2_B_g$g)
plot_data2_B_g$Class <- as.factor(plot_data2_B_g$Class)

plot_data3_B <- read.csv("./ex09_plot_data3_B.csv")
plot_data3_B <- plot_data3_B[, -(4:6)]
plot_data3_B_g <- gather(plot_data3_B, key = Class, value = FBOA, BOA_diffuse1, 
                           BOA_direct1, BOA_diffuse2, BOA_direct2)
plot_data3_B_g$omega <- as.factor(plot_data3_B_g$omega)
plot_data3_B_g$g <- as.factor(plot_data3_B_g$g)
plot_data3_B_g$Class <- as.factor(plot_data3_B_g$Class)
plot_data3_B_g$theta <- as.factor(plot_data3_B_g$theta)
  
```

### Layer Number Dependency (Fig. 1)

First of all the dependence of the model output on the thickness of each cloud layer $(\Delta\tau)$ is tested. Therefore the cloud optical thickness $\tau$ is fixed to $10$ and the number of layers in this cloud is varied. The result shows that from roughly layer 25 onward there is no real difference in the model output anymore. To allow for uncertainties a layer number of $50$ for $\tau=10$ is deemed appropriate. This in turn leads to a $\Delta\tau$ of $0.2$. This value is used as a default for the model. The number of layer for each cloud is then easily calculated depending on its overall optical thickness: $$n_{layer}=\tau / \Delta\tau$$

```{r Plotting, echo=FALSE, fig.cap="Layer Number Dependency of the Irradiance"}
library("ggplot2")
library("latex2exp")
delta_tau_depend <- ggplot(data = delta_tau_layer_depend_g)+
  geom_point(aes(n_layer, Irradiance, color = Class))+
  geom_line(aes(n_layer, Irradiance, color = Class))+
  labs(title="Layer Number Dependency of the Irradiance")+
  ylab("Percent")+
  xlab("Number of Layers")
print(delta_tau_depend)

```

### Photon Number Dependency (Fig. 2)

To see what amount of photons have to be run through the model a simple test is made. The results show no visible difference between $5,000$ and $20,000$ photons. Therefore the number of photons is set to: $$n_{photon}=5000$$

```{r echo=FALSE, fig.cap="Photon Number Dependency of the Irradiance"}
photon_depend <- ggplot(data = plot_data1_g)+
  geom_point(aes(n_photon, Irradiance, color = Class))+
  labs(title="Photon Number Dependency of the Irradiance")+
  ylab("Percent")+
  xlab("Number of Photons")
print(photon_depend)
```

### Task A (Fig. 3)

The following graph is facetted and shows the Dependency of the Irradiance of the optical thickness $\tau$ with respect to the single scattering albedo $\tilde\omega \, (0.9, 0.99, 1)$ and the asymmetry parameter $g \, (-0.9 - 0.9)$.  
The absorbed fraction of photons is highest at a low $\tilde\omega$ and a high $g$ (bottom left of graph), which is in line with the explanations from exercise 8. With a decrease in $\tilde\omega$ and $g$ more radiation is scattered forward and can be increasingly absorbed.  
An increase in $g$ also leads to more $F^\downarrow_{BOA}$ which is countered by an higher optical thickness of the cloud. A more negative $g$, meaning mostly backward scattering, has an opposite effekt leading to increased $F^\uparrow_{TOA}$ which is enforced by higher values of $\tau$. 
In general it can be said that an in crease in $\tau$ leads to more $F^\uparrow_{TOA}$ and less $F^\downarrow_{BOA}$.

```{r echo=FALSE, fig.cap="Optical Thickness Dependency of the Irradiance with Respect to Single Scattering Albedo and Asymmertry Parameter"}
all_depend <- ggplot(data = plot_data2_B_g)+
  geom_point(aes(tau, Irradiance, color = Class))+
  geom_line(aes(tau, Irradiance, color = Class))+
  facet_grid(g~omega)+
  labs(title="Optical Thickness Dependency of the Irradiance with Respect to \n 
       Single Scattering Albedo and Asymmertry Parameter")+
  ylab("Percent")+
  xlab("Optical Thickness")
print(all_depend)
```

### Task B (Fig. 4)

Now the diffuse fraction  $f_{diff}=F^\downarrow_{BOA,diff}/F^\downarrow_{BOA}$ is investigated. As can be seen the two methods for differentiating $F^\downarrow_{BOA}$ differ strongly for optically thin clouds $(\tau < 5)$ but show the same for $\tau > 5$. This is probably due to the lack of precision in method 2 for it classifies all $F^\downarrow_{BOA}$ as diffuse (BOA_diffuse2).  
But from method 1 it can be said that a increase of $\tau$ and of the solar zenith angle $\theta_0$ leads to an increase in $f_{diff}$.  
Higher values of the asymmetry parameter $g$ seem to lead to an higher $f_{diff}$.

```{r echo=FALSE, fig.cap="Diffuse and Direct Fraction of Downward Irradiance in Dependence of Optical Thickness, Solar Zenith Angle and Asymmetry Parameter"}
fboa_plot <- ggplot(data = plot_data3_B_g)+
  geom_point(aes(tau, FBOA, color = Class))+
  geom_line(aes(tau, FBOA, color = Class))+
  facet_grid(g~theta)+
  labs(title = "Diffuse and Direct Fraction of Downward Irradiance in Dependence of\n Optical Thickness, Solar Zenith Angle and Asymmetry Parameter")+
  ylab("Percent")+
  xlab("Optical Thickness")
print(fboa_plot)
```
